GMSE: a general tool for management strategy evaluation

### 2017-09-15 notebook/presentation/GMSE_presentation.Rmd
Nearing the endish now... finally

---
title: "GMSE: a general tool for management strategy evaluation"
subtitle: "https://bradduthie.shinyapps.io/GMSE_presentation/#1"
author: "Brad Duthie"
date: "19 September 2017"
output: 
  ioslides_presentation:
    logo: Confoobio-stacked-logo.png
    css: slide.css
    widescreen: no
    fig_width: 5
    fig_height: 5
runtime: shiny
---

@@

## Conservation conflict

@@

## Background on ConFooBio

Resolving conflicts between food security and biodiversity conservation under uncertainty

@@

- What leads to conflict
- How to solve conflict
- Impact on sustainability

All of the above are characterised by uncertainty

@@

[Adami et al. (2016) *Phys. Life. Rev.* **19**:1-26](http://dx.doi.org/10.1016/j.plrev.2016.08.015)

@@

**Images:** (1) [Public license](https://commons.wikimedia.org/wiki/File:US_and_USSR_nuclear_stockpiles.svg), (2) [Wilkinson 1990](https://projects.ncsu.edu/project/evoresources/Evolutionary%20medicine/Unitsofselection%20DM%20copy/wilkinson-foodsharing.pdf), (3) [Isaacdm](https://commons.wikimedia.org/wiki/File:PhysicsofPenalty.JPG?uselang=en-gb), (4) [Public license](http://www.publicdomainpictures.net/view-image.php?image=180560&picture=brexit-referendum-uk)

@@

## Simple game-theoretic models insufficient

@@

- pay-offs are known and fixed
- Actions are clear and always available
- Players behave rationally
- Game rules are known to all
- There is a clear rank order of outcomes

## Genetic algorithms

Show the basic idea
Use the gamesGA package to illustrate it in a game-theoretic context

@@

@@

## Introduce GMSE

@@

![](GMSE_Logo_Goose_300.png)

@@

@@

## Introduce GMSE

@@

![](GMSE_Logo_Goose_300.png)

- Written in R and C
- Requires R (>= 3.3.3)
- License GPL (>=2)

@@

>>>>

---
title: "GMSE: a general tool for management strategy evaluation"
subtitle: "https://bradduthie.shinyapps.io/GMSE_presentation/#1"
author: "Brad Duthie (@bradduthie)"
date: "19 September 2017"
output: 
  ioslides_presentation:
    logo: Confoobio-stacked-logo.png
    css: slide.css
    widescreen: no
    fig_width: 5
    fig_height: 5
runtime: shiny
---

@@

<!--- ![](African_Bush_Elephant.jpg) --->

![](grouse_shooting.jpg)

@@

**Image:** [Public](https://commons.wikimedia.org/wiki/File:Hardy_Heywood_-_The_Grouse_Shoot_1898.jpg) [Domain](https://commons.wikimedia.org/wiki/File:Henharrier.jpg)
<!--- **Image:** [Muhammad Mahdi Karim](https://en.wikipedia.org/wiki/User:Muhammad_Mahdi_Karim) --->

<!--- ![](African_Bush_Elephant.jpg) --->

![](grouse_shooting2.jpg)

**Image:** [Public](https://commons.wikimedia.org/wiki/File:Hardy_Heywood_-_The_Grouse_Shoot_1898.jpg) [Domain](https://commons.wikimedia.org/wiki/File:Henharrier.jpg)
<!--- **Image:** [Muhammad Mahdi Karim](https://en.wikipedia.org/wiki/User:Muhammad_Mahdi_Karim) --->

## The ConFooBio Project

**Resolving conflicts between food security and biodiversity conservation under uncertainty**

@@

![](group.png)

## The ConFooBio Project

**Resolving conflicts between food security and biodiversity conservation under uncertainty**

- What leads to conservation conflict
- How to resolve conservation conflict
- Impact on sustainability of social-ecological systems

@@

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; [Adami et al. (2016) *Phys. Life. Rev.* **19**:1-26](http://dx.doi.org/10.1016/j.plrev.2016.08.015)

@@

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; **Images:** (1) [Public license](https://commons.wikimedia.org/wiki/File:US_and_USSR_nuclear_stockpiles.svg), (2) [Wilkinson 1990](https://projects.ncsu.edu/project/evoresources/Evolutionary%20medicine/Unitsofselection%20DM%20copy/wilkinson-foodsharing.pdf), (3) [Isaacdm](https://commons.wikimedia.org/wiki/File:PhysicsofPenalty.JPG?uselang=en-gb), (4) [Public license](http://www.publicdomainpictures.net/view-image.php?image=180560&picture=brexit-referendum-uk)

@@

## Stake-holders are complex

**Simple models assume that stake-holders:**

- Know what pay-offs will be
- Have clearly defined choices
- Behave rationally
- Understand the games
- Prioritise outcomes clearly

@@

## Conservation conflicts are complex {.smaller}

 - **Asymmetric payoffs**: Payoffs of choices may depend instead on stakeholder identity.

![](payoff_matrix_asym.png)

## But most games are more complex {.smaller}

 - **Asymmetric payoffs**: Payoffs of choices may depend instead on stakeholder identity.
  - **Multiple choices**: Players might have three or more choices.

![](payoff_matrix_cho.png)

## But most games are more complex {.smaller}

 - **Asymmetric payoffs**: Payoffs of choices may depend instead on stakeholder identity.
 - **Multiple choices**: Players might have three or more choices.
 - **Multiple players**: More than two stakeholders might interact at one time.

![](payoff_matrix_3pl.png)

## But most games are more complex {.smaller}

 - **Asymmetric payoffs**: Payoffs of choices may depend instead on stakeholder identity.
 - **Multiple choices**: Players might have three or more choices.
 - **Multiple players**: More than two stakeholders might interact at one time.
 - **Iterated interactions**: Strategies maybe affected by interaction history.

```{r, echo=FALSE}
rounds  <- as.character(1:9);
players <- c("Player 1","Player 2");
plays   <- sample(x=c(0,1), size=18, replace=TRUE);
plays[plays == 0] <- "A";
plays[plays == 1] <- "B";
history <- matrix(data=plays, nrow=2);
rownames(history) <- players;
colnames(history) <- rounds;
kable(noquote(history));
```

![](payoff_matrix_rd10_sm.png)

##{.smaller}

![](gmse_overview_sm_ga.png)

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; [Bunnefeld et al. (2011) *Trends. Ecol. Evol.* **26**:441-447](https://doi.org/10.1016/j.tree.2011.05.003)

![](gmse_overview_sm_ga2.png)

@@

![](gmse_overview_sm_ga2.png)

@@

## Genearlised MSE {.smaller}

@@

@@

![](GMSE_Logo_Goose_375.png)

@@

## Genearlised MSE ('GMSE' R package)

@@

- Individual-based
- Spatially explicit
- Written in R and C
- Requires R (>= 3.3.3)
- License GPL (>=2)
- *v0.2.2.7* [available](https://cran.r-project.org/package=GMSE) on CRAN
- Browser-based GUI [available](https://bradduthie.shinyapps.io/gmse_gui/) via shiny
- Source code [available](https://github.com/bradduthie/gmse) on GitHub
- Intended for a wide range of software users

![](GMSE_Logo_Goose_375.png)

@@

## GMSE: Natural resources

- Spatially explicit landscape
- Discrete individual resources
- Individual birth, movement, death

![](sep_land.gif)

## GMSE: Resource observation

- Four types of observation options
    1. Counting on a subset of landscape
    2. Mark and recapture
    3. Linear transect sampling
    4. Block transect sampling
- Mechanistic sampling process
- Generates simulated data frame used by the manager

![](binos.png)

**Image:** [Public Domain](https://commons.wikimedia.org/wiki/File:Fernglas(alt).JPG)

## GMSE: Management

- One manager analyses observations and compares with a target abundance
- Five possible policy options (scaring, culling, castration, feeding, helping)
- Fixed budget and a costs for setting policy
- Genetic algorithm anticipates policy effects from previous actions to set costs of actions for stakeholders

![](balance.png)

## GMSE: Stake-holder actions {.smaller}

- Any number of stake-holders
- Stake-holders want to maximise harvest of resources or minimise resource effects on their land
- Seven possible actions (scaring, culling, castration, feeding, helping, tend crops, kill crops)
- Fixed budget for performing actions of potentially differing costs
- Genetic algorithm anticipates effects of actions on harvest or land yield

![](harvesting.jpg)

**Image:** [Public Domain](https://commons.wikimedia.org/wiki/File:Libiaz-zniwa1938.a.jpg)

## Running the gmse() function {.smaller}

```{r, echo = TRUE}
sim <- gmse(land_ownership = TRUE, stakeholders = 10, observe_type = 1,
            res_death_K = 2000, manage_target = 1000, RESOURCE_ini = 1000,
            user_budget = 5000, manager_budget = 5000, res_consume = 1,
            scaring = TRUE, fixed_mark = 50, fixed_recapt = 300,
            plotting = FALSE);
```

## Explain two types of plots

```{r, echo = FALSE, fig.height = 6, fig.width = 7}
plot_gmse_results(res = sim$resource, obs = sim$observation, land = sim$land,
                  agents = sim$agents, paras = sim$paras, ACTION = sim$action,
                  COST = sim$cost);
```

## Other plot type
